### 2012-07-01 doc/acknowledgments.tex
Started reconstruction paper

TJC acknowledges support from STFC in the form of a research studentship.
%
PJM was given support by the TABASGO and Kavli foundations, and the Royal 
Society, in the form of research fellowships.
%
MWA \ldots. SH \ldots. SHS \ldots.
%
TT acknowledges support from the NSF through CAREER award NSF-0642621,
and from the Packard Foundation through a Packard Research Fellowship.
% 
LVEK acknowledges the support by an NWO-VIDI programme subsidy
(programme number 639.042.505).
%
This research is supported in part by the National Science Foundation under
Grant No. PHY99-07949.