### 2012-07-01 doc/addresses.tex

\def\ioa{Institute of Astronomy, University of Cambridge,
  Madingley Rd, Cambridge, CB3 0HA, UK}
\def\oxford{Dept.\ of Physics, University of Oxford, 
  Keble Road, Oxford, OX1 3RH, UK}
\def\kipac{Kavli Institute for Particle Astrophysics and Cosmology, 
 Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA}
\def\ucsb{Dept.\ of Physics, University of California, 
  Santa Barbara, CA 93106, USA}
\def\davis{Dept.\ of Physics, U.C.~Davis, Davis, CA 95616, USA}
\def\kapteyn{Kapteyn Astronomical Institute, University of Groningen, 
  P.O.Box 800, 9700 AV Groningen, The Netherlands}

\def\collettemail{\tt t.collett@ast.cam.ac.uk}

\def\packard{Packard Research Fellow}