### 2012-10-02 doc/references.tex
draft update

\begin{thebibliography}{99}

\bibitem[\protect\citeauthoryear{Auger et al.}{2009}]{Auger++2009} 
Auger M.~W., Treu T., Bolton A.~S., Gavazzi R., Koopmans L.~V.~E., Marshall 
P.~J., Bundy K., Moustakas L.~A., 2009, ApJ, 705, 1099

\bibitem[\protect\citeauthoryear{Auger et al.}{2010}]{Auger++2010} 
Auger M.~W., Treu T., Bolton A.~S., Gavazzi R., Koopmans L.~V.~E., Marshall 
P.~J., Moustakas L.~A., Burles S., 2010, ApJ, 724, 511

\bibitem[\protect\citeauthoryear{Baltz, Marshall, 
\& Oguri}{2009}]{BMO} Baltz E.~A., Marshall P., Oguri M., 2009, JCAP, 1, 15

\bibitem[\protect\citeauthoryear{Behroozi, Conroy, 
\& Wechsler}{2010}]{Behroozi++2009} Behroozi P.~S., Conroy C., Wechsler R.~H., 2010, ApJ, 717, 379

\bibitem[\protect\citeauthoryear{Ben{\'{\i}}tez}{2000}]{BPZ} Ben{\'{\i}}tez N., 2000, ApJ, 536, 571

\bibitem[\protect\citeauthoryear{Collett et 
al.}{2012}]{Collett++2012a} Collett T.~E., Auger M.~W., Belokurov V., 
Marshall P.~J., Hall A.~C., 2012, MNRAS, 424, 2864

\bibitem[\protect\citeauthoryear{De Lucia 
\& Blaizot}{2007}]{deLucia+Blaizot2007} De Lucia G., Blaizot J., 2007, MNRAS, 375, 2

\bibitem[\protect\citeauthoryear{Falco, Gorenstein, 
\& Shapiro}{1985}]{Falco++1985} Falco E.~E., Gorenstein M.~V., Shapiro I.~I., 1985, ApJ, 289, L1

\bibitem[\protect\citeauthoryear{Fassnacht, Koopmans, 
\& Wong}{2011}]{Fassnacht++2011} Fassnacht C.~D., Koopmans L.~V.~E., Wong K.~C., 2011, MNRAS, 410, 2167

\bibitem[\protect\citeauthoryear{Gavazzi et 
al.}{2008}]{Gavazzi++2008} Gavazzi R., Treu T., Koopmans L.~V.~E., 
Bolton A.~S., Moustakas L.~A., Burles S., Marshall P.~J., 2008, ApJ, 677, 
1046

\bibitem[\protect\citeauthoryear{Hilbert et 
al.}{2009}]{2009A&A...499...31H} Hilbert S., Hartlap J., White S.~D.~M., Schneider P., 2009, A\&A, 499, 31

\bibitem[\protect\citeauthoryear{Holz 
\& Wald}{1998}]{Holz+Wald1998} Holz D.~E., Wald R.~M., 1998, PhRvD, 58, 063501

\bibitem[\protect\citeauthoryear{Keeton 
\& Moustakas}{2009}]{Keeton+Moustakas2009} Keeton C.~R., Moustakas L.~A., 2009, ApJ, 699, 1720

\bibitem[\protect\citeauthoryear{Linder 
\& Holz}{2004}]{Linder+Holz2004} Linder E.~V., Holz D.~E., 2004, AAS, 36, \#131.02

\bibitem[\protect\citeauthoryear{Nandra, Lasenby, 
\& Hobson}{2012}]{Nandra++2012} Nandra R., Lasenby A.~N., Hobson M.~P., 2012, MNRAS, 422, 2945

\bibitem[\protect\citeauthoryear{Navarro, Frenk, 
\& White}{1997}]{NFW} Navarro J.~F., Frenk C.~S., White S.~D.~M., 1997, ApJ, 490, 493

\bibitem[\protect\citeauthoryear{Neto et al.}{2007}]{Neto2007} 
Neto A.~F., et al., 2007, MNRAS, 381, 1450

\bibitem[\protect\citeauthoryear{Oguri 
\& Marshall}{2010}]{Oguri+Marshall2010} Oguri M., Marshall P.~J., 2010, MNRAS, 405, 2579

\bibitem[\protect\citeauthoryear{Oke}{1974}]{Oke1974} Oke 
J.~B., 1974, ApJS, 27, 21

\bibitem[\protect\citeauthoryear{Springel et 
al.}{2005}]{Springel++2005} Springel V., et al., 2005, Natur, 435, 629

\bibitem[\protect\citeauthoryear{Suyu}{2012}]{Suyu2012b} Suyu 
S.~H., 2012, arXiv, arXiv:1202.0287

\bibitem[\protect\citeauthoryear{Suyu et al.}{2012}]{Suyu++2012a} 
Suyu S.~H., et al., 2012, arXiv, arXiv:1208.6010

\bibitem[\protect\citeauthoryear{Suyu et al.}{2010}]{Suyu++2010} 
Suyu S.~H., Marshall P.~J., Auger M.~W., Hilbert S., Blandford R.~D., 
Koopmans L.~V.~E., Fassnacht C.~D., Treu T., 2010, ApJ, 711, 201

\bibitem[\protect\citeauthoryear{Vale 
\& White}{2003}]{Vale+White2003} Vale C., White M., 2003, ApJ, 592, 699

\bibitem[\protect\citeauthoryear{Wong et al.}{2011}]{Wong++2011} 
Wong K.~C., Keeton C.~R., Williams K.~A., Momcheva I.~G., Zabludoff A.~I., 
2011, ApJ, 726, 84

\bibitem[\protect\citeauthoryear{\tom{which paper}}{{9999}}]{xxx} 
None

\end{thebibliography}