### 2012-10-30 doc/references.tex
more intro in hilbert 2011 etc, MS section added

\begin{thebibliography}{99}

\bibitem[{{Auger} {et~al.}(2007){Auger}, {Fassnacht}, {Abrahamse}, {Lubin}, \&
  {Squires}}]{AugerEtal2007}
{Auger}, M.~W., {Fassnacht}, C.~D., {Abrahamse}, A.~L., {Lubin}, L.~M., \&
  {Squires}, G.~K. 2007, \aj, 134, 668

\bibitem[\protect\citeauthoryear{Auger et al.}{2009}]{AugerEtal2009} 
Auger M.~W., Treu T., Bolton A.~S., Gavazzi R., Koopmans L.~V.~E., Marshall 
P.~J., Bundy K., Moustakas L.~A., 2009, ApJ, 705, 1099

\bibitem[\protect\citeauthoryear{Auger et al.}{2010}]{AugerEtal2010} 
Auger M.~W., Treu T., Bolton A.~S., Gavazzi R., Koopmans L.~V.~E., Marshall 
P.~J., Moustakas L.~A., Burles S., 2010, ApJ, 724, 511

\bibitem[\protect\citeauthoryear{Baltz, Marshall, 
\& Oguri}{2009}]{BMO} Baltz E.~A., Marshall P., Oguri M., 2009, JCAP, 1, 15

\bibitem[\protect\citeauthoryear{Behroozi, Conroy, 
\& Wechsler}{2010}]{BehrooziEtal2009} Behroozi P.~S., Conroy C., Wechsler R.~H., 2010, ApJ, 717, 379

\bibitem[\protect\citeauthoryear{Ben{\'{\i}}tez}{2000}]{BPZ} Ben{\'{\i}}tez N., 2000, ApJ, 536, 571

\bibitem[{{Brada{\v c}} {et~al.}(2009){Brada{\v c}}, {Treu}, {Applegate},
  {Gonzalez}, {Clowe}, {Forman}, {Jones}, {Marshall}, {Schneider}, \&
  {Zaritsky}}]{BradacEtal2009}
{Brada{\v c}}, M., {et~al.} 2009, \apj, 706, 1201

\bibitem[Cropper et al.(2012)]{EUCLID} Cropper, M., Cole, R., 
James, A., et al.\ 2012, arXiv:1208.3369

\bibitem[\protect\citeauthoryear{Collett et 
al.}{2012}]{CollettEtal2012a} Collett T.~E., Auger M.~W., Belokurov V., 
Marshall P.~J., Hall A.~C., 2012, MNRAS, 424, 2864

% \bibitem[Dalal(2005)]{Dalal2005} Dalal, N.\ 2005, 25 Years After 
% the Discovery: Some Current Topics on Lensed QSOs,

\bibitem[{{Dalal} {et~al.}(2005){Dalal}, {Hennawi}, \& {Bode}}]{DalalEtal2005}
{Dalal}, N., {Hennawi}, J.~F., \& {Bode}, P. 2005, \apj, 622, 99

\bibitem[\protect\citeauthoryear{De Lucia 
\& Blaizot}{2007}]{deLucia+Blaizot2007} De Lucia G., Blaizot J., 2007, MNRAS, 375, 2

\bibitem[Fadely et al.(2010)]{FadelyEtal2009} Fadely, R., Keeton, 
C.~R., Nakajima, R., \& Bernstein, G.~M.\ 2010, \apj, 711, 246

\bibitem[\protect\citeauthoryear{Falco, Gorenstein, 
\& Shapiro}{1985}]{FalcoEtal1985} Falco E.~E., Gorenstein M.~V., Shapiro I.~I., 1985, ApJ, 289, L1

\bibitem[{{Fassnacht} \& {Lubin}(2002)}]{Fassnacht+Lubin2002}
{Fassnacht}, C.~D., \& {Lubin}, L.~M. 2002, \aj, 123, 627

\bibitem[\protect\citeauthoryear{Fassnacht, Koopmans, 
\& Wong}{2011}]{FassnachtEtal2011} Fassnacht C.~D., Koopmans L.~V.~E., Wong K.~C., 2011, MNRAS, 410, 2167

\bibitem[\protect\citeauthoryear{Gavazzi et 
al.}{2008}]{GavazziEtal2008} Gavazzi R., Treu T., Koopmans L.~V.~E., 
Bolton A.~S., Moustakas L.~A., Burles S., Marshall P.~J., 2008, ApJ, 677, 
1046

\bibitem[Gunnarsson et al.(2006)]{GunnarssonEtal2006} Gunnarsson, C., 
Dahl{\'e}n, T., Goobar, A., J{\"o}nsson, J., M{\"o}rtsell, E.\ 2006, \apj, 640, 417

\bibitem[{{Hilbert} {et~al.}(2007){Hilbert}, {White}, {Hartlap}, \&
  {Schneider}}]{HilbertEtal2007}
{Hilbert}, S., {White}, S.~D.~M., {Hartlap}, J., \& {Schneider}, P. 2007,
  \mnras, 382, 121

\bibitem[\protect\citeauthoryear{Hilbert et 
al.}{2009}]{HilbertEtal2009} Hilbert S., Hartlap J., White S.~D.~M., Schneider P., 2009, A\&A, 499, 31

\bibitem[Hilbert et al.(2011)]{HilbertEtal2011} Hilbert, S., Gair, 
J.~R., \& King, L.~J.\ 2011, \mnras, 412, 1023

\bibitem[Holder 
\& Schechter(2003)]{Holder+Schechter2003} Holder, G.~P., \& Schechter, P.~L.\ 2003, \apj, 589, 688

\bibitem[\protect\citeauthoryear{Holz 
\& Wald}{1998}]{Holz+Wald1998} Holz D.~E., Wald R.~M., 1998, PhRvD, 58, 063501

\bibitem[Holz 
\& Linder(2005)]{Holz+Linder2005} Holz, D.~E., \& Linder, E.~V.\ 2005, \apj, 631, 678

\bibitem[{{Keeton} \& {Zabludoff}(2004)}]{Keeton+Zabludoff2004}
{Keeton}, C.~R., \& {Zabludoff}, A.~I. 2004, \apj, 612, 660

\bibitem[\protect\citeauthoryear{Keeton 
\& Moustakas}{2009}]{Keeton+Moustakas2009} Keeton C.~R., Moustakas L.~A., 2009, ApJ, 699, 1720

% \bibitem[\protect\citeauthoryear{Linder 
% \& Holz}{2004}]{Linder+Holz2004} Linder E.~V., Holz D.~E., 2004, AAS, 36, \#131.02

\bibitem[LSST Science Collaboration et al.(2009)]{LSST} LSST 
Science Collaboration, Abell, P.~A., Allison, J., et al.\ 2009, 
arXiv:0912.0201

\bibitem[Macci{\`o} et al.(2008)]{MaccioEtal2008} Macci{\`o}, A.~V., 
Dutton, A.~A., \& van den Bosch, F.~C.\ 2008, \mnras, 391, 1940

\bibitem[{{Momcheva} {et~al.}(2006){Momcheva}, {Williams}, {Keeton}, \&
  {Zabludoff}}]{MomchevaEtal2006}
{Momcheva}, I., {Williams}, K., {Keeton}, C., \& {Zabludoff}, A. 2006, \apj,
  641, 169

\bibitem[Nakajima et al.(2009)]{NakajimaEtal2009} Nakajima, R., 
Bernstein, G.~M., Fadely, R., Keeton, C.~R., 
\& Schrabback, T.\ 2009, \apj, 697, 1793

\bibitem[\protect\citeauthoryear{Navarro, Frenk, 
\& White}{1997}]{NFW} Navarro J.~F., Frenk C.~S., White S.~D.~M., 1997, ApJ, 490, 493

\bibitem[\protect\citeauthoryear{Neto et al.}{2007}]{Neto2007} 
Neto A.~F., et al., 2007, MNRAS, 381, 1450

\bibitem[{{Oguri} \& {Takahashi}(2006)}]{Oguri+Takahashi2006}
{Oguri}, M., \& {Takahashi}, K. 2006, \prd, 73, 123002

\bibitem[\protect\citeauthoryear{Oguri 
\& Marshall}{2010}]{Oguri+Marshall2010} Oguri M., Marshall P.~J., 2010, MNRAS, 405, 2579

\bibitem[\protect\citeauthoryear{Oke}{1974}]{Oke1974} Oke 
J.~B., 1974, ApJS, 27, 21

\bibitem[\protect\citeauthoryear{Springel et 
al.}{2005}]{SpringelEtal2005} Springel V., et al., 2005, Natur, 435, 629

\bibitem[\protect\citeauthoryear{Suyu}{2012}]{Suyu2012a} Suyu 
S.~H., 2012, arXiv, arXiv:1202.0287

\bibitem[\protect\citeauthoryear{Suyu et al.}{2012}]{SuyuEtal2012} 
Suyu S.~H., et al., 2012, arXiv, arXiv:1208.6010

\bibitem[\protect\citeauthoryear{Suyu et al.}{2010}]{SuyuEtal2010} 
Suyu S.~H., Marshall P.~J., Auger M.~W., Hilbert S., Blandford R.~D., 
Koopmans L.~V.~E., Fassnacht C.~D., Treu T., 2010, ApJ, 711, 201

\bibitem[\protect\citeauthoryear{Vale 
\& White}{2003}]{Vale+White2003} Vale C., White M., 2003, ApJ, 592, 699

\bibitem[{{Wang} \& {Dai}(2011)}]{Wang+Dai2011}
{Wang}, F.~Y., \& {Dai}, Z.~G. 2011, \aap, 536, A96

\bibitem[{{Williams} {et~al.}(2006){Williams}, {Momcheva}, {Keeton},
  {Zabludoff}, \& {Leh{\'a}r}}]{WilliamsEtal2006}
{Williams}, K.~A., {Momcheva}, I., {Keeton}, C.~R., {Zabludoff}, A.~I., \&
  {Leh{\'a}r}, J. 2006, \apj, 646, 85

\bibitem[\protect\citeauthoryear{Wong et al.}{2011}]{WongEtal2011} 
Wong K.~C., Keeton C.~R., Williams K.~A., Momcheva I.~G., Zabludoff A.~I., 
2011, ApJ, 726, 84

\bibitem[{{Wong} {et~al.}(2012){Wong}, {Ammons}, {Keeton}, \&
  {Zabludoff}}]{WongEtal2012}
{Wong}, K.~C., {Ammons}, S.~M., {Keeton}, C.~R., \& {Zabludoff}, A.~I. 2012,
  \apj, 752, 104

% \bibitem[\protect\citeauthoryear{\tom{which paper}}{{9999}}]{xxx} 
% None

\end{thebibliography}