### 2012-11-16 doc/acknowledgments.tex
replace accidentally deleted files

TEC acknowledges support from STFC in the form of a research studentship.
%
PJM was given support by the Kavli Foundation and the Royal 
Society, in the form of research fellowships.
%
MWA \ldots
%
SH \ldots
%
SHS, ZG  and TT gratefully acknowledge support from the Packard Foundation in the form of a Packard Research Fellowship to TT and from the National Science Foundation grant AST-0642621.
% 
LVEK acknowledges the support by an NWO-VIDI programme subsidy
(programme number 639.042.505).
%
This research is supported in part by the National Science Foundation under
Grant No. PHY99-07949.