### 2012-12-06 doc/paper-lcr.tex
finalized draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Pangloss: Testing the reconstruction method
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[useAMS,usenatbib,a4paper]{mn2e}
%% letterpaper
%% a4paper

%\voffset=-0.6in

% Packages:
\input psfig.sty
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}

% Macros:
\input{macros.tex}
\input{addresses.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[Line of Sight Mass Reconstruction]
{Reconstructing the Lensing Mass in the Universe \\
from Photometric Catalogue Data}

\author[Collett \etal]{%
  Thomas~E.~Collett$^{1}$\thanks{\collettemail},
  Philip~J.~Marshall$^{2}$,
  Matthew~W.~Auger$^{1}$,
  Stefan~Hilbert$^{3}$,
\newauthor{%
  Sherry~H.~Suyu$^{3,4}$,
  Zachary~Greene$^{4}$,
  Tommaso~Treu$^{4}$\thanks{\packard},
  Christopher~D.~Fassnacht$^{5}$,}
\newauthor{%
  L\'eon~V.~E.~Koopmans$^{6}$,
  Maru\v{s}a Brada\v{c}$^{5}$,
  Roger~D.~Blandford$^{3}$} 
  \medskip\\
  $^1$\ioa\\
  $^2$\oxford\\
  $^3$\kipac\\
  $^4$\ucsb\\
  $^5$\davis\\
  $^6$\kapteyn
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\date{to be submitted to MNRAS}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}\pubyear{2012}

\maketitle

\label{firstpage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}

High precision cosmological distance measurements towards individual objects
such as time delay gravitational lenses or type Ia supernovae are affected by
weak lensing perturbations by galaxies and groups along the line of sight. In
time delay gravitational lenses, ``external convergence,'' $\kappax$, can
dominate the uncertainty in the inferred distances. In this paper we attempt
to reconstruct $\kappax$, due to line of sight structure using a simple halo
model.
%
We use mock catalogues from the Millennium Simulation, and calibrate and 
compare our reconstructed $\pr(\kappax|\data)$ to ray-traced $\kappax$ truth
values:  taking into account realistic uncertainties on redshift and stellar
masses, \tom{we find that the reconstruction of $\kappax$ provides an improvement in
precision of $\sim$50\% over previously published schemes.} 
%
We find that the lowest $\kappax$ lines of sight have the best constrained 
$\pr(\kappax)$. In anticipation of future samples with thousands of lenses,
we find that selecting the third of the systems with the
highest precision $\kappax$ estimates gives a  sub-sample of unbiased time
delay distance measurements \phil{with (on average) just} 1\% uncertainty due
to line of sight external convergence effects. Photometric data alone is
sufficient to pre-select the best-constrained lines of sight, and can be done
before investment in lightcurve monitoring. 
%
\phil{Conversely, we show that somehow selecting lines of sight
with high external shear could, with the reconstruction model presented here,
induce biases of up to 1\% in time delay distance.}  We suggest areas for the
improvement of this general analysis framework (including more sophisticated
treatment of high mass structures) that should allow yet more accurate
cosmological inferences to be made.

\end{abstract}

% Full list of options at http://www.journals.uchicago.edu/ApJ/instruct.key.html

\begin{keywords}
  gravitational lensing   --
  methods: statistical    --
  galaxies: halos         --
  galaxies: mass function  --
  cosmology: observations
\end{keywords}

\setcounter{footnote}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}

Every distant object we observe has had its apparent shape distorted,
and size and total brightness magnified (or demagnified) by a compound
weak gravitational lens constructed from all the mass distributed
between us and it. As \citet{Vale+White2003} and \citet{HilbertEtal2007}
showed, there are no empty lines of sight through our universe. This
fact makes gravitational lensing a potentially important source of
systematic error for any estimate of luminosity (or distance); this 
issue has been 
raised for \eg type Ia supernovae by
\citet[][]{Holz+Wald1998,Holz+Linder2005}, for gamma ray bursts by
\citet[][]{Oguri+Takahashi2006,Wang+Dai2011},  and for high redshift
galaxies by \citet{WyitheEtal2011}, among others.

Along lines of sight containing strong gravitational lenses, the perturbative
effects of line of sight mass structure have been found to be particularly
important, with foreground and background structures having a significant
effect on the inferred lensing cross-section \citep[\eg][]{WongEtal2012} and
distance ratios \citep[][]{DalalEtal2005}. Indeed, \citet{SuyuEtal2010} found
that in time delay lens cosmography the so-called ``external convergence''
$\kappax$ due to mass structures along the (unusually over-dense) line of
sight to the quadruply-imaged radio source B1608$+$656 had to be included in
their analysis, and was the dominant source of uncertainty in their 5\%
measurement of the time  delay distance.  Large samples of time-delay lenses
are expected to be discovered in the next decade in ground-based optical
imaging surveys \citep{Oguri+Marshall2010};  the external convergence will
have to be understood increasingly well in order to prevent its correction
from dominating the systematic error budget.

While it is rare for three galaxies to line up well enough for both of
the background sources to be strongly lensed \citep{GavazziEtal2008,CollettEtal2012a}, 
the large size of dark matter halos makes partial alignments -- such that they
act as perturbing weak lenses --  a near certainty. The large scale of the
perturbers means the external lensing perturbations can be approximated
by a quadrupole lens characterised only by external convergence and shear.
The external shear, $\gammax$ can potentially be recovered in the
mass-modelling of the lens, but the convergence is undetectable from the
image positions, shapes and relative fluxes; this is the well-known {\emph{ 
mass-sheet degeneracy}} \citep[see e.g.][for details]{FalcoEtal1985}.

Since the mass-sheet degeneracy prevents any estimate of $\kappax$ from the 
strong lens modelling, additional information is required. Weak lensing measurements
can been used for rich lines of sight \citep{NakajimaEtal2009, FadelyEtal2009}, but 
for lines of sight with low galaxy density we must attempt to reconstruct the distribution
of mass along the line of sight. Attempts to reconstruct the mass distribution in strong
lens fields have focused on understanding the external shear which seems to be large in most strong lenses.
Surveys by \citet{Fassnacht+Lubin2002,AugerEtal2007,WilliamsEtal2006,MomchevaEtal2006}
all found groups of galaxies hosting or near to known gravitational lenses.
\citet{WongEtal2011} estimated $\Pr(\gammax)$ given the data of \citet{WilliamsEtal2006}
and \citet{MomchevaEtal2006} but found significant discrepancy between the predicted shear
distribution and the external shear demanded by the strong lens model. 
\citeauthor{WongEtal2011} also found that both line of
sight structures and the group of galaxies in each lens plane contribute
significant proportions of the shear.

For lines of sight without strong lenses, magnification is the more relevant
quantity than convergence; \citet{GunnarssonEtal2006} used a galaxy halo model
combined with simple scaling relations to reconstruct supernovae lines of
sight and found that the dispersion in apparent source brightness due to
lensing magnification could be reduced by a factor of two. Conversely
\citet{KarpenkaEtal2012} used the brightness dispersion in type Ia supernovae
to infer the parameters of both their galaxy halo model and mass-to-light
scaling relation. Both \citeauthor{KarpenkaEtal2012} and
\citet{JonssonEtal2010} were able to  detect lensing effects in the supernova
data this way.

Large numerical simulations can also be used to estimate global convergence
distributions. \citet{Holder+Schechter2003} and \citet{DalalEtal2005} carried
out ray-tracing calculations in N-body simulations to estimate the
distribution of external shear values.  \citet{HilbertEtal2009} performed
similar ray-tracing experiments through the Millennium Simulation
\citep{SpringelEtal2005}, generating a predicted $\kappax$ at every position
in a simulated sky. \citet{HilbertEtal2009} found that after removing the
matter on the strong lensing plane \MS lines of sight with strong lenses were
not biased towards high $\kappax$, although selection functions for
discovering samples of strong lenses were not taken into account.

The \MS results have been used to analyse real observations as well. 
\citet{SuyuEtal2010} selected \MS lines of sight by their apparent
galaxy over-density, in a 45 arcsec radius aperture down to $i < 24.5$,
to match the observed over-density towards the time delay lens
B1608$+$656 \citep{FassnachtEtal2011}.  The resulting distribution of
$\kappax$ values from the ray-tracing was taken to be an estimate of 
$\Pr(\kappax)$, which was then marginalised over when inferring the time
delay distance in this system. A similar procedure was used in \citet{SuyuEtal2012}.
In a companion paper to this one, Greene et al. (submitted) investigate
improvements to this method by weighting the galaxy counts by galaxy
mass, and perpendicular distance to the line of sight, again using the
\MS mock catalogues and their associated $\kappax$ values to construct
$\Pr(\kappax|\data)$. For the most over-dense lines of sight their
results  show a $\sim$20\% improvement over
number counts alone, but little improvement for less dense ones.

In this paper, we combine the halo model reconstruction approach of
\citeauthor{GunnarssonEtal2006}, \citeauthor{JonssonEtal2010}, 
\citeauthor{WongEtal2011} and others, with the
idea of calibrating to simulations from \citeauthor{SuyuEtal2010}
Taking the $\kappax$ values from
\citeauthor{HilbertEtal2009}'s ray-tracing calculations as the ``truth''
that we have to recover, we use the \MS mock galaxy catalogues first to
{\it calibrate} the reconstruction, and account for unseen mass and
voids, and then to {\it test} the accuracy of the line of sight mass
reconstruction under this assumption. Probabilistically assigning mass
and redshift to every observed galaxy in a given observed field, we
generate Monte Carlo sample line of sight mass distributions, and so
construct the PDF $\Pr(\kappax|\data)$ for that field. We then emulate
the combination of many such PDFs to quantify the residual bias that
would be translated to the global (including cosmological) parameters, 
\phil{if left unaccounted for.}
In doing so, we aim to answer the following questions:

\begin{itemize}

\item Faint galaxies, filaments and dark structures will not appear in
any photometric object catalogue, but they will contribute convergence at
some level. How much of the total line of sight 
convergence comes from visible
galaxies, and how much effect do dark structures and voids have?

\item Can the true line of sight convergence be recovered from a calibrated
halo model reconstruction? What scatter and residual bias are induced by the
reconstruction process, and how might these be reduced in the future?

\item Both the detectability of lenses, and the selection of sub-samples of
them for further observation, could potentially induce a selection bias in
external convergence. How should we select objects to achieve the highest
accuracy convergence reconstruction? Can such a selection be made robustly,
using the reconstruction results?

\end{itemize}

This paper is organized as follows. We review the relevant gravitational
lensing theory in~\Sref{sec:theory}, and the Millennium Simulation
ray-traced convergences and mock catalogues in \Sref{sec:MS}, before
introducing our simple reconstruction  model in \Sref{sec:model}. We
then test this model in two phases: first, in \Sref{sec:knownMh+z}, with
known redshift and halo mass for every galaxy in a lens field, in order
to quantify the irreducible uncertainty due to unseen mass, and second,
in \Sref{sec:obsMstar+z}, with realistic observational uncertainties on
the observed galaxies' stellar masses and redshifts. In
\Sref{sec:biases} we investigate the potential systematic error induced
by selecting a subset of lines of sight. We discuss our results in
\Sref{sec:discuss} before concluding in \Sref{sec:conclude}.

Throughout this paper magnitudes are given in the AB system and
we adopt the Millennium Simulation's ``concordance'' parameters for our reference cosmology, \ie
$h=0.73$, $\Omega_m=0.25$ and $\Omega_\Lambda=0.75$, where the symbols indicate
the Hubble Constant in units of 100 km s$^{-1}$ Mpc$^{-1}$ and the matter and
dark energy density of the Universe in units of the critical density.
%\citep[e.g.\ ][]{KomEtal09}.

\comments{ When inferring global quantities (such as the cosmological
parameters), combining the results from a large number of independent objects
will tend to reduce the uncertainty due to external convergence
\citep[\eg][]{Holz+Linder2005}, unless the lines of sight are not
representative of the global population. If the line of sight selection
function is not correctly modelled some residual systematic error will remain
even as the statistical uncertainty decreases. A selection function may be
introduced either at the object detection stage, or later on when making
decisions about which objects to study further. Large samples of lenses are
expected to be discovered in the next decade in ground-based optical imaging
surveys \citep{Oguri+Marshall2010}: both the detectability of these lenses, 
and the selection of sub-samples of them for further observation, could be
sensitive to the external convergence.

Several attempts have been made to do better than simply constructing a large
sample of objects, and averaging over the resulting convergence distribution.
The weak lensing effect can be detected observationally  by measuring the
small distortions it induces on the images of many background galaxies in the
field \citep[see e.g.][for a review]{Schneider2006}.  \citet{NakajimaEtal2009}
used deep HST imaging to infer $\kappax=0.17\pm0.06$ from weak lensing
measurements; this measurement was used in the time delay distance measurement
of \citet{FadelyEtal2009}. A precise weak lensing estimate of $\kappax$ can
only be made if the number density of of weakly lensed, measurable galaxies is
high. }

\comments{In order to propagate the effect of line-of-sight perturbations into 
uncertainties on time-delay distances we require the probability density function (PDF) for the external convergence
given our available data $\data$, $\Pr(\xkappa|\data)$}

% PHIL's REVIEWS OF THE LITERATURE
\comments{\citet{GunnarssonEtal2006} investigated a related PDF,  for the
magnification, using a galaxy halo model. Applying empirical galaxy
scaling relations both to simulate mock catalogues and then to reconstruct
the line of sight mass distribution. While they did not explore highly
over-dense lines of sight, or the impact of groups and clusters, they
found that under their simplifying assumptions, the dispersion in
apparent source brightness due to lensing magnification could be reduced
by a factor of two.}
%
\comments{\citet{WongEtal2011} estimated the PDF for the external shear,
$\Pr(\gammax|\data)$ from their survey data, running similar Monte Carlo
reconstructions of the mass in nine lens fields based on
\citeauthor{WilliamsEtal2006} and \citeauthor{MomchevaEtal2006}'s
photometric and spectroscopic measurements of galaxies close to the line
of sight. They then  compared the resulting predicted shear
distributions with the external shear demanded by the strong lens model.
Their reconstructions showed most of the shear to have been generated by
bright galaxies within 2~arcminutes of the lens, and that both line of
sight structures and the group of galaxies in each lens plane contribute
significant proportions of the shear. They found significant
discrepancies between the lens model shear and the Monte Carlo
predictions, but were unable to distinguish between the environment
modelling and the strong lens modelling as the cause of these
discrepancies.}
%
\comments{\citet{KarpenkaEtal2012} used a halo model to predict deterministically 
the magnification due to
galaxies along the line of sight to type Ia supernovae, and inferred the
parameters of scaling relations between light and mass in these galaxies
from hundreds of supernova lightcone simultaneously. Following
\citet{JonssonEtal2010}, they subtract off the mean convergence to ensure that
the universal mass budget is balanced, and compare models where the halos are
all have either truncated singular isothermal profiles, or 
NFW profiles. Both groups were able to 
detect lensing effects in the supernova data this way.}

\begin{figure*}
\includegraphics[width=\textwidth]{figs/viewofalightcone.eps}
\caption[magcut]{Four different views of a slightly over-dense \MS line
of sight, with a $\kappax$ of 0.03. 
{\it Top row, left:} The positions of galaxies projected on the sky. The area
of the circles is proportional to observed $i$-band flux; the brightest
object shown has an $i$ magnitude of 17.7. 
{\it Top row, centre:} The angular sizes of halos projected on the sky.
Red and blue regions lie within the NFW scale radius and virial radius
of each halo, respectively. There are essentially no empty lightcones. 
{\it Top row, right:} The individual $\kappax$ contributions of
each halo, assuming a \citet{BMO} truncated NFW 
profile and the \citet{Neto2007}
mass-concentration relation. Comparison of
this panel and the centre panel illustrates the relative importance of
proximity to the line of sight.
{\it Bottom:} A view along the redshift axis, showing only halos with
$|x|<0.3$ arcmin. The area of the points is proportional to each halo's
mass: the most massive halo shown has $1.6\times10^{12}\Msun$. The
optical axis is shown by the dashed line, while the dotted lines mark
the lens and source planes for a B1608-like strong lens.}
\label{fig:lightcone}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Theoretical Background}
\label{sec:theory}

\comments{
Convergence, or Ricci focusing, occurs when a gravitational lens focuses
the rays in a given bundle. This focusing can cause distant objects to
appear brighter, and larger than they would if the lens were removed.
The convergence from an isolated mass sheet is the ratio of the projected surface
mass density ($\Sigma$) \devided by the
critical surface mass density at the redshift of the mass sheet
($\Sigma_{\rm cr}$),
\be
\kappax= \frac {\Sigma}
                              {\Sigma_{\rm cr}(\zd,\zs)}
\ee
where 
\be 
\label{eq:sigcrit} 
\Sigma_{\rm cr}(\zd,\zs) \equiv \frac{c^2 D_{\rm os}}{4 \pi G D_{\rm od} D_{\rm ds}}.
\ee
and the $D$'s are the angular diameter distances between the objects
referred to in the subscripts: o is the observer, d is the deflector and
s is the source. If the surface mass density of the sheet exceeds
$\Sigma_{\rm cr}$, multiple images of the source can occur, otherwise
only one image will be observed, although that image will still be
perturbed relative to the unlensed case.

Strong lensing occurs when a massive object and background source are almost perfectly
aligned along a line of sight. The light from the background source is deflected by the
lens galaxy; this deflection allows multiple images of the background source to form
at stationary points of the time delay function. For an isolated lens, the time delay
function can be calculated from
\be \label{eq:T} 
\Delta t(\bmath{\theta},\bmath{\beta}) = \frac {1}{c} \frac{D_{\rm od} D_{\rm os}}{D_{\rm ds}} (1+z_{\rm d})\, \phi(\bmath{\theta},\bmath{\beta}),
\ee
where $\bmath{\theta}$ is the observed source position, $\bmath{\beta}$ is the 
unlensed source position, $z_{\rm d}$ is the redshift of the lens, $\phi(\bmath{\theta},\bmath{\beta})$ is
the Fermat potential. The Fermat potential is given by
\be \label{eq:FP}
\phi(\bmath{\theta},\bmath{\beta})\equiv \left[\frac{(\bmath{\theta}-\bmath{\beta})^2}{2}-\psi(\bmath{\theta}) \right], 
\ee
where $\psi(\bmath{\theta})$ is the lens potential, derived from the projected dimensionless
surface mass density, $\kappa(\bmath{\theta})$, by 
\be \label{eq:psikappa}
\kappa(\bmath{\theta})=\frac{1}{2}\nabla^2\psi(\bmath{\theta}).
\ee
}

In strong lens systems where the source is time variable, the images do not 
vary simultaneously; the optical path length for each image is different due
to relativistic and geometric effects. The difference in optical path length
causes a time-delay between the light curves of each image.
\citet{FalcoEtal1985} showed that the presence of additional matter (in the
form of a mass-sheet) along the line of sight has no observable effect except
to rescale the time-delays. As such, it is necessary to include $\kappax$ in
the lens modelling, if cosmological parameters are to be estimated accurately
and precisely from observed time delays \citep{SuyuEtal2010}.

If there is external convergence present that is not included in the
lens modelling, then the time delay distance -- inferred assuming $\kappax
= 0$ -- will be $(1-\kappax)$ more than the true value of $\Dt$:
\be 
\label{eq:MassSheet:Dtbias}
\Dt^{\rm{true}}=\frac{\Dt^{{\kappax = 0}}}{1-\kappax}.
\ee

We can hence estimate the true distance {\it only} if we have additional
knowledge of $\kappax$. Since $\kappax$ is typically small the absolute
uncertainty on the estimate of $\kappax$ corresponds to the fractional
uncertainty with which time-delay distances can be \infered.

Mass distributions that are physically associated with the lens galaxy will
affect the stellar dynamics of the lens galaxy, so that dynamical observations
can break the internal mass-sheet degeneracy by providing an additional
estimate of the lens galaxy's mass
\citep[e.g.,][]{Koopmans2004,SuyuEtal2010}.  Our aim is to quantify the
lensing effect of massive objects, regardless of their position along the line
of sight. \phil{However, in this paper we restrict ourselves to reconstructing
the line of sight portion of the mass distribution in any given field, which
allows us to work in the weak lensing regime. Treatment of sightlines
containing strong lensing events, and the impact of physically-associated mass
distributions on both the strong lensing and stellar dynamical effects are
postponed to further work.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The Millennium Simulation}
\label{sec:MS}

In order to test the accuracy of our convergence estimates, we need to
know the true convergence for each line of sight. We cannot use  real
lines of sight for this,  but must use simulated lines of sight instead.
In this section we briefly review the Millennium Simulation, the ray
tracing calculations that have been carried out in it, and the mock
galaxy catalogues that have been produced.

The \MS \citep{SpringelEtal2005} is a cosmological N-body simulation of
Dark Matter structures in a cubic region approximately 680~Mpc in
co-moving size, followed from redshift 127 to the present day. With an
approximate halo mass resolution of $2\times10^{10}{\rm M}_{\odot}$
(corresponding approximately to a galaxy with luminosity $0.1L^{*}$), it
provides a detailed prediction for the distribution of dark structures
present in the Universe, under the assumptions of the $\Lambda$CDM model
of hierarchical structure formation, cosmological densities as given at
the end of \Sref{sec:intro}, and $\sigma_8 = 0.9$.
Galaxy properties,
such as stellar masses, luminosities and colours, were assigned to the
simulated halos according to a semi-analytic model for galaxy formation
\citep{DeLucia+Blaizot2007}: the resulting predictions for the galaxy
luminosity function and correlation function match the observational
data very well.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Convergence from Ray Tracing}
\label{sec:MS:raytracing}

\citet{HilbertEtal2009} calculated the lensing convergence using the
``lightcone'' dark matter only  output of the \MS, providing 1.7 degree
square  mock sky maps of this quantity. In this process, the dark matter
density from the simulation was projected onto a set of lens planes at
discrete redshifts, adaptively gridded and smoothed, and then the second
derivatives of the  lensing potential required for the components of the
magnification matrix were computed using a multiple lens plane ray tracing
algorithm. The first order approximation to this algorithm \citep[equation 17
of][]{HilbertEtal2009} gives the total convergence at a given sky
position~$\bmath{\theta}$ as the simple summation of the surface density in
each lens plane, weighted by the inverse critical surface density $\Sigma_{\rm
cr}$ for that plane's redshift:
\be
\kappax(\bmath{\theta}) = \sum_i \frac{\Sigma_i(\bmath{\theta})}
                                     {\Sigma_{\rm cr}(z_i,\zs)}.
\ee
\citet{HilbertEtal2009} found that this approximation is accurate to a few percent or better
in $\kappax$, even on 30 arcsecond scales.

To define a test line of sight, we draw a random sky position from
within the convergence maps, bi-linearly interpolate between the pixels
of the map, and store the result as the ``true'' convergence for that
line of sight, $\kappax^{\rm true}$.

\comment{PJM: This means our sightlines are not typical of strong lenses, because they
don't include the local overdense region typically found near massive lens
galaxies. Our lines of sight have typical mass distributions for a lens line
of sight {\it after excision of the lens plane}. I have tried to explain
this above.}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Mock Photometric Galaxy Catalogues}
\label{sec:MS:mocks}

We construct mock photometric galaxy catalogues by selecting all \MS galaxies
within a circular aperture of a given angular radius centred on each randomly
selected line of sight position, with apparent $i$-band magnitude brighter
than a given limit. The parent catalogue, from \citet{HilbertEtal2011},
contains galaxy positions and magnitudes, all of which include lensing effects
(deflections and magnifications). We also have access to the underlying galaxy
halo masses and redshifts, which we will use to calibrate the reconstruction,
and also to explore any sources of bias and scatter. Since the true
convergence is that of only the dark matter, we focus on reconstructing the
dark matter halos alone. Our model for doing this is outlined in the next
section.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Estimating Convergence: The Halo Model Approximation}
\label{sec:model}

We require a method for estimating the convergence at any point on the sky,
given a catalogue of observed galaxy properties such as positions,
brightnesses, colours, and possibly stellar masses and redshifts, for every
galaxy in the field of the lens, down to some magnitude limit. Since we cannot
observe the surface mass density of each halo directly, we need some way of
estimating the mass of each halo in the catalogue, so that we can compute its
contribution to the total $\kappax$ along the line of sight \citep[as in
\eg][]{GunnarssonEtal2006,KarpenkaEtal2012}.  This mass
assignment recipe will be uncertain, and also incomplete, but it can be
calibrated with cosmological simulations, which represent a significant
additional source of information to what is present in the data.

Transforming an observed photometric galaxy catalogue into a  catalogue of
halo masses, positions, and redshifts will enable us to attempt a
reconstruction of the convergence induced by every halo near a line of sight. 
We also need to account for the convergence due to dark structures and the
divergence due to voids; we do not model voids or dark structures, but their
effects are included statistically by calibrating the convergence in
halos, $\kappah$, against the ray-traced convergence, $\kappax$, for \MS lines
of sight.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Halos}
\label{sec:model:halos}

Cosmological dark matter simulations have shown that dark matter
halos are well-approximated by NFW profiles 
\citep{NFW}. We assume each halo to have a spherical mass distribution with 
density given by
\be\label{eq:rhonfw}
\rho(r) = 
\frac{\rho_0}{\left(r/r_{s}\right)
\left(1+r/r_{s}\right)^{2}}.
\ee
Here, $r_{s}$ is a characteristic scale 
radius of the cluster, representing the point where
the density slope transitions from $r^{-1}$ to $r^{-3}$. This radius is
related to the virial radius of the halo by $r_{s}~=~r_{200}/c$, where $c$ is
the concentration parameter, which can be estimated from the halo's mass,
using a mass--concentration relation. Typically more massive halos are less
concentrated, but there is some scatter; we use the relation of
\citet{Neto2007} to estimate $c$ from the mass enclosed within $r_{200}$,
which we denote as $M_{200}$. We find our results do not change if the
\citet{MaccioEtal2008} mass--concentration relation is used instead.

If one integrates the density profile of an NFW profile out to infinite
radius, the total mass diverges. Similarly, if the universe is
homogeneously populated with NFW halos, the projected surface mass along
any line of sight will also be divergent.\footnote{At large radius,
$\Sigma_{\rm nfw} \propto R^{-2}$, but the differential number of halos
centred within an annulus of width $\dee R$ is given by $\dee N_{\rm
annulus} \propto R \dee R$, so  $\Sigma_{\rm total} \propto
\int_{0}^{\infty} R^{-1} \dee R$, which diverges logarithmically} Since
infinite mass is unphysical, the profile must be truncated at some
point. Several truncation profiles have been suggested
\citep[e.g][]{BMO}, but beyond several virial radii, the amount of
matter associated with a halo is likely to be low. In this work we
assume the truncated NFW profile
\be\label{eq:bmoprofile}
\rho(r) = 
\frac{\rho_{\rm NFW} (r)}{1+\left(r/r_{t}\right)^2},
\ee
which is the same as the NFW profile in the limit that the truncation
radius, $r_t$ goes to infinity; the shear and convergence from such a
profile are derived in \citet{BMO}.
 We use a truncation radius of five times
the virial radius, but our results are robust for any choice of $r_t>2
\times r_{200}$.

Many studies have shown that galaxies have total mass density profiles that
are approximately isothermal in their inner regions
\citep[\eg][]{AugerEtal2010} due to the more concentrated stellar mass
component.
This has been confirmed by galaxy-galaxy weak lensing measurements 
\citep[\eg][]{Mandelbaum,Leauthaud}. We note that the two halo term that
dominates the galaxy-galaxy lensing signal at large radii is explicitly taken
into account in our model, where every galaxy is assigned a dark matter halo.
At large radii an NFW-like profile is probably appropriate, but may not be
correct when a halo is very close to the line of sight.

Each halo in our catalogue contributes to
the line of sight convergence by
\be
\label{eq:kappai}
\kappa_i =\Sigma_{i}/\Sigma_{\mathrm{cr}}(z_i,\zs),
\ee
where $\Sigma_{\mathrm{cr}}(z_i,\zs)$ is the critical surface density \comments{was defined in \Eref{eq:sigcrit}}.
Following the first order approximation of \citet{HilbertEtal2009}
outlined in \Sref{sec:MS} above, we compute 
the total convergence from all the halos along the line
of sight using
\be 
\label{eq:kappasummu}
\kappah = \sum_{i} \kappa_i.
\ee

To estimate a halo's contribution to $\kappah$ we must first know the
halo's mass. For the lines of sight of the \MS, halo masses are
known, but for  real lines of sight the halo masses have to be \infered
from whatever data is available. 
In this work we investigate the use of the empirical stellar mass --
halo mass relation for this purpose. The primary advantage of this
approach is that it is a simple, ``one size fits all'' relation that can
be applied regardless of the stellar mass observed. We discuss
additional observables that could be used to improve the precision of
the halo mass inference in \Sref{sec:discuss} below.

We use the relation of \citet{BehrooziEtal2010} to infer halo masses
from stellar masses; the details of this \proceedure are outlined in 
\Aref{appendix:MSMH}. Given an uncertain stellar mass and redshift for
each galaxy in the  catalogue, we draw a sample halo mass from the PDF
that describes this uncertainty (\Eref{eq:mhalo-mstar}) and use
\Eref{eq:kappai} to compute its contribution to the convergence,
$\kappa_i$; this can be done for all the halos along a specific line of sight
and summed to give a sample value of $\kappah$;  repeatedly applying
this \proceedure allows us to build up a histogram of $\kappah$ values
consistent with the data, and hence characterize $\pr(\kappah|\data)$.

This PDF contains a hidden assumption of an uninformative prior PDF for
$\kappah$ -- in the case of infinitely poorly measured stellar masses and
redshifts, $\pr(\kappah|\data)$ will have very long tails corresponding to
very over-dense lines of sight that we do not believe exist. The resolution of this
is to divide out the effective prior PDF that was applied during the halo mass
estimation process (which is broad, and hence close to uniform), and apply an
additional prior on $\kappah$, given by the global underlying $\kappah$
distribution from the simulation. In the limiting  case of very poor
photometric data, $\pr(\kappah|\data)$ then defaults to this distribution, as
required.

% \todo{All}{Discuss this. It ties us to the simulation, but only weakly. Is the
% argument about poor data valid? Tom?}

% TEC: yes it's valid. It only really ties us to the halo mass function of the
% simulation. It's probably tied at the same level as the "global P(\kappa) as
% derived from Stefan's ray-tracing"

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Accounting for voids, filaments and other dark structures}
\label{sec:model:voids}

Our halo model accounts for the convergence contribution of density
perturbations that can be associated with light. Filaments and dark
substructures contribute additional convergence, while voids contribute a
divergence; these structures must be included to produce an unbiased estimate
of $\kappax$. In particular, neglecting voids would lead to a heavily biased
estimate of $\kappax$, since the halo model's $\kappah$ can only be positive.
In principle the absence of galaxies implies something about the presence of a
void, but we do not currently attempt to use this information. In general, it
is difficult to account for the unseen mass, since we do not have a good model
for its density structure.

The solution we present here is to calibrate the relationship between the halo
convergence $\kappah$ and the overall convergence $\kappax$ (as would be
calculated by ray tracing through the full density field)  using the
simulations themselves. The halo modelling procedure outlined in
\Sref{sec:model:halos} allows  us to estimate  the convergence due to halos
given observations of  mass, redshift and position,
$\pr(\kappah|\data)$, but what we are interested in is the total
convergence along the line of sight $\pr(\kappax|\data)$. We can obtain
this by considering the expression: 
\begin{equation} 
\Pr(\kappax|\data) = 
   \int \dee\kappah  \Pr(\kappax|\kappah,\data) \Pr(\kappah|\data)
   \label{eq:kappaconv}
\end{equation} 
The first term in the integrand relates the convergence due to model halos,
$\kappah$, to the true convergence, $\kappax$. This conditional distribution
can be constructed from the \MS catalogues, by computing
$\kappah$ from the true halo masses and redshifts 
along each selected line of sight, and then
accumulating $\kappah$ and $\kappax$
along a large number of such sightlines. 
For any given line of sight we can then estimate
$\pr(\kappah|\data)$ as described above, and then multiply it by
$\Pr(\kappax|\kappah)$ and integrate out the intermediate parameter
$\kappah$.

\phil{However, if the conversion between stellar mass and halo mass is very
uncertain, $\pr(\kappah|\data)$ is shifted relative to what it would
have been given perfect knowledge of the halo masses.  This is because the
conversion from stellar mass to halo mass and into $\kappah$ is highly
asymmetric: a long tail of high halo masses gives rise to a tendency to
overestimate $\kappah$. If this shift is ignored, the resulting
$\Pr(\kappax|\data)$ will be systematically biased towards high values of
$\kappax$. Instead, we need the  distribution of $\kappax$ from all
calibration lines of sight that have a $\pr(\kappah|\data)$ that is identical
to the $\pr(\kappah|\data)$ for the real line of sight {\it given the same
data quality}; however, we found this approach unfeasible  given our finite
number of calibration lines of sight. A working compromise is to use 
the median of the $\pr(\kappah|\data)$ distributions in the calibration.}

We  take a large number of
calibration lines of sight from the \MS catalogues and generate a mock
stellar mass catalogue for each one. We then reconstruct each lightcone in the
same way as we would an observed field,
estimate $\pr(\kappah|\data)$, and extract its median. Accumulating
these median values $\kappah^{\rm med}$ and 
their corresponding true values of $\kappax$, we can form the 
conditional distribution $\pr(\kappax|\kappah^{\rm med})$. 
Then, to infer $\kappax$ from an observed photometric catalogue, we 
estimate $\pr(\kappah|\data)$ 
using the halo model, compute the median value 
$\kappa_{\rm h,obs}^{\rm med}$, and then use a modified
version of 
\Eref{eq:kappaconv} to infer $\kappax$: 
\bea
\Pr(\kappax|\kappah^{\rm med},\data) &=& \int \dee\kappah^{\rm med} 
   \Pr(\kappax|\kappah^{\rm med},\data) \Pr(\kappah^{\rm med}|\data) \notag \\
\label{eq:calkappaconv}   
\eea
\phil{This integral is trivial, since $\Pr(\kappah^{\rm med}|\data)$ is
a delta function centred on the median of the inferred PDF for $\kappah$.
However,}
because we only have a finite number of calibration lines of sight, we use all
calibration lines of sight with $\kappa_{\rm h,cal}^{\rm med} =
\kappah^{\rm med}\pm0.003$ to form $\Pr(\kappax|\kappah^{\rm
med},\data)$.
\comments{, with each calibration sightline weighted equally. PJM: why is
this? Surely one should weight by the value of the Gaussian function implied
by that ``$\pm 0.003$''?}

This procedure of reducing the full PDF for $\kappah$ to its median increases
the importance of the realism of the simulation: the simulation needs to be as
similar as possible to the real universe or there is a possibility of
systematic errors. Conversely, making use of all the information
in the photometric catalogue yields $\kappax$ estimates of increased
precision.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Results: Perfect Halo Data}
\label{sec:knownMh+z}

We now put the reconstruction \proceedure outlined above into practice, first
inferring $\kappax$ from $\kappah$ in the case of hypothetical galaxy
catalogues with noiseless redshifts and halo masses. The reason for doing this
is to check the validity of the calibration process, and then also to 
investigate the primary sources of the convergence. It will also provide us
with a measure of the intrinsic uncertainty introduced by our assumption of
all illuminated halos being spherically-symmetric truncated NFW halos with
Neto et al concentration--mass relation.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{How is $\kappah$ related to $\kappax$?}

\Fref{fig:jointkh-k} shows the conditional distribution of $\kappax$ given 
$\kappah$, derived from $10^5$ randomly selected \MS lines of sight
and assuming noise-free halo masses and redshifts. 
\phil{(In this case, $\pr(\kappah|\data)$ is almost a delta function since the only uncertainty is halo
concentration, which has minimal effect on $\pr(\kappah|\data)$.  
)}.
% In this case, the PDF : we can read
% off  $\pr(\kappax|\data)$ just by taking a thin vertical slice through this 
% distribution. 
% Note that the joint PDF plotted in this figure is the product
% of  the conditional distribution that we need for calibration, and also the
% prior PDF for $\kappah$ that we need to ensure that the global distribution of
% $\kappah$ values is defaulted to in the case of poor observational data. 
At fixed $\kappah$ we find in \Fref{fig:jointkh-k} that the scatter in 
$\kappax$
grows with $\kappah$; our reconstruction is better at reproducing under-dense
lines of sight than over-dense lines of sight. The effect of ignoring voids
and the smooth mass component is evident in this plot: $\kappah$ is
significantly higher than $\kappax$, at any given $\kappax$ value. 
\phil{Conditioning on $\kappah$ gives a PDF for
$\kappax$ centred at the correct place, by construction.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[width=\columnwidth]{figs/cornerplot.eps}
\caption[Biased?]{The conditional distribution of 
$\kappax$ given $\kappah$, in the case where  
we have perfect knowledge of halo mass and redshift.
$10^5$ reconstructed lines of
sight were used to make this plot. 
Solid (dashed) lines enclose 68\% (95\%) of the conditional probability.
$\kappah$ traces $\kappax$, but with a positive offset which arises from
$\kappah$ not accounting for any voids or dark structures.
At fixed $\kappah$ the scatter in $\kappax$ grows rapidly with $\kappah$.
\tom{The thin diagonal line follows $\kappax=\kappah$.}}
\label{fig:jointkh-k}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{How accurate is the $\kappah - \kappax$ calibration?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[width=\columnwidth]{figs/globaldist.eps}
\caption[magcut]{Recovering the global $\pr(\kappax)$ (shown in red). The
black curve shows a histogram of  inferred  $\kappax$ values from $10^{5}$
reconstructed lines of sight.  The reconstructions were performed given
perfect knowledge of halo mass and redshift -- in this case the reconstruction
recovers the correct convergence.}
\label{fig:globaldist}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\includegraphics[width=\textwidth]{figs/where_is_the_kappa.eps}
\caption[magcut]{Which objects dominate the external convergence for a line of
sight? From left to right, the three figures show the cumulative contribution
to the total convergence from individual halos ($\kappah$) as a function of
1) distance from the line of sight, 2) magnitude and 3) redshift.   In each
panel, the solid line is the median cumulative contribution over a large
sample of lines of sight, while dashed and dot-dashed show the ranges enclosing 68
and 95 percent of the lines of sight respectively.}
\label{fig:where}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Let us define the ``bias'' of the reconstruction of a given line of sight as
the difference between the expectation value of $\kappax$ over its posterior
PDF $\pr(\kappax|\data)$ and the known true value $\kappax^{\mathrm{true}}$.
Let us also define the ``width,'' $\sigma_{\kappa}$, to be half the width of
the interval containing the central $68\%$ of the posterior probability
$\pr(\kappax|\data)$.

In an ensemble of $10^{5}$ reconstructed lines of sight, we find the mean bias
to be $-1\times 10^{-5}$, and the mean width to be 0.01. This is approximately
1.8 times smaller than the width of the global $\pr(\kappax)$.  We find that
our inferred $\kappax$ PDFs are consistent with the global $\kappax$
distribution; \Fref{fig:globaldist} shows that sampling from the
$\pr(\kappax|\data)$ inferred from reconstructions of  $10^{5}$ lines of
sight  produces a distribution of $\kappax$ values over the sky identical to
the global $\pr(\kappax)$. Given perfect knowledge of halo mass and redshift
then, the calibration \proceedure provides an unbiased estimate of $\kappax$
that is $1.8$ times as precise as using the global $\pr(\kappax)$. 
\phil{This is the maximum precision we can obtain using the current calibrated
halo model.}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Which halos dominate the $\kappah$ distribution?}

Before investigating the impact of imperfect knowledge of halo
mass and redshift, we investigate the uncertainties induced by limits on the
magnitude of the observed galaxy sample, and on the field of view. The halos
in our catalogue are populated by galaxies and given magnitudes according to
the semi-analytic model of \citet{DeLucia+Blaizot2007}: by applying magnitude
cuts to our catalogue, we can investigate the amount of scatter caused by
unobserved halos.

\Fref{fig:where} shows the cumulative contribution to $\kappah$ as a function
of projected distance from the line of sight, magnitude and redshift. We find
that most of the convergence comes from halos close to the line of sight. 
Over half of the  convergence typically comes from halos within 30 arcseconds
of the line of sight; by 2 arcminutes, this fraction is 85\%. The contribution
from halos beyond 2 arcminutes is relatively constant at a contribution of
0.008$\pm$0.003. We find that ignoring halos beyond 2 arcminutes has no effect
on the precision of the reconstruction; this is shown in \Fref{fig:radcut}. As
a function of magnitude, we find that $\kappah$ is dominated by objects with
magnitudes between $i=18$ and $i=24$. Objects brighter than $i=18$ are either
too rare, or too close to the observer, to make a significant contribution to
the convergence. Objects fainter than $i=24$ are too small to be important,
unless they are extremely close to the line of sight, but we find that
including halos fainter than $i=24$ does not improve the reconstruction. 
\Fref{fig:magcut} shows the scatter on $\kappah-\kappax$ (where $\kappah$ is
shifted so that its mean is zero, \phil{to emulate the effect of the eventual
calibration}) as a function of reconstruction magnitude limit. We do not
expect a deeper survey to decrease the size of the uncertainty in mapping
$\kappah$ onto $\kappax$. Halos at all redshifts out to the source redshift
(1.4, here) contribute to the convergence, but the largest contribution comes
from halos with $z \sim z_{\rm source}/2$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[width=\columnwidth]{figs/radius_scatter.eps}
\caption[magcut]{The 16, 50 and 84th percentiles of $\kappah$ minus
$\kappax$ as a function of the limiting radius of the halo
reconstruction. $\kappah$ has been shifted such that
$\left\langle\kappah\right\rangle=0$ \phil{to emulate the effect of the
eventual calibration}. 
The majority of the constraining power
comes from reconstructing halos within 2~arcmin of the line of sight.}
\label{fig:radcut}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[width=\columnwidth]{figs/mag_scatter.eps}
\caption[magcut]{The 16, 50 and 84th percentiles of $\kappah$ minus
$\kappax$ as a function of the limiting $i$ band depth of the halo
reconstruction. $\kappah$ has been shifted such that
$\left\langle\kappah\right\rangle=0$ \phil{to emulate the effect of the
eventual calibration}. The majority of the constraining power
comes from reconstructing halos with magnitudes between $18<i<24$.}
\label{fig:magcut}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Testing the Halo Model Reconstruction on Mock Galaxy Catalogues}
\label{sec:obsMstar+z}

We now move on to consider the halo model reconstruction of line-of-sight mass
distributions given noisy astronomical observables. A typical imaging survey
can be expected to provide measurements of the positions and magnitudes of
galaxies in a field;  spectra for some of the objects may either come from a
synergistic survey, or from targeted follow-up. In this section we quantify
the uncertainties induced by inferring the halo mass and redshift from these
observables.

%\flag{Should we show the relevant calibration plots, analogous to Figure 2?
%Seems a bit odd to give Figure 2 but not illustrate the {\it actual
%calibration procedure we follow on real data}... Maybe we could make one that
%had the conditional distribution for each data quality?}

%TEC: I think that could lead to figure overload; the concept is given throughout
% the paper, and it's well illustrated by figure 2.

As described in \Sref{sec:model:halos}, in this work we attempt to  infer halo
masses from measurements of stellar mass. We will investigate two main sources
of uncertainty: the stellar masses themselves, and placing halos at
photometric redshifts. Much work has already focused on using photometric
colours to infer stellar mass \citep[\eg][]{AugerEtal2009} and redshifts
\citep[\eg][]{BPZ}: we will estimate the likely uncertainties on stellar mass
and redshift based on this work.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Making Mock Observational Catalogues}

\phil{While the \MS catalogues do already contain stellar masses for each
galaxy, we do not use them for two reasons. The first is that at the low mass
end, the dark matter-only \MS satellite galaxy halos are highly  stripped
relative to what we might expect in a universe containing baryons, leading to
a mismatch between the observed and simulated stellar mass--halo mass
relations.  The second is that we wanted to be able to perform the functional
test of trying to recover the convergence having assumed the correct stellar
mass--halo mass relation.} For these reasons, we assign a new true stellar
mass to each halo in the \MS catalogues, according to the empirical stellar
mass--halo mass relation of \citet{BehrooziEtal2010}. From these we simulate
observed stellar masses by drawing samples from  $\pr(\log{M^{*}_{\rm
obs}}|\log{M^{*}_{\rm true}})$ which we take to be a Gaussian of width
$\sigma_{M_*}$ centred on $\log(M*_{\mathrm {true}})$. Where a spectroscopic
redshift exists, stellar masses can be estimated with typical uncertainties of
0.15 dex \citep{AugerEtal2009}; however with photometric redshifts stellar
mass uncertainties are typically three times as large. We use
$\sigma_{M_*}=0.15$ dex for halos with a spectroscopic redshift and
$\sigma_{M_*}=0.45$ dex otherwise. For photometric redshift uncertainties we
draw a redshift from $\pr(z_{\rm true}|z_{\rm obs})$ which we take to be a
Gaussian of width $0.1(1+z_{\rm spec})$ centred on $z_{\rm spec}$, where
$z_{\rm spec}$ is the halo's true redshift in the \MS catalogue.

\comment{The following is relevant to all sections on inferring $\kappax$
given imperfect data, so it belongs here and not in the spectroscopic redshift
first subsection.}

Given an uncertain stellar mass and redshift it is possible to infer a
halo mass using the stellar mass--halo mass relation of
\citet{BehrooziEtal2010}. This \proceedure requires an inversion of the
relation given in \citet{BehrooziEtal2010} and correctly inverting the
relation's uncertainties requires care: the \proceedure we use to do this
is given in \Aref{appendix:MSMH}. By drawing sample halo masses
and redshifts, we can infer a sample $\kappah$ using the \proceedure of
\Sref{sec:model}. Repeatedly drawing samples allows us to
estimate $\pr(\kappah|\data)$ for each reconstructed line of sight. We then
transform this into our target PDF $\pr(\kappax|\data)$ as described above.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Reconstructing $\kappax$ given a spectroscopic redshift for every object}

%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[width=\columnwidth]{figs/Width.eps}
\caption{The widths of the \infered PDFs $\pr(\kappah|\data)$ for
$10^5$ lines of sight, given different quality of data. 
Blue: spectroscopic redshift for every halo with $i<26$; 
Red: spectroscopic redshift for every halo with $i<23$, 
and every halo with $i<24$ within 1 arcminute, while all other objects just
have photometric redshifts;
Green: all the objects in the field have only photometric redshifts. The
vertical dot-dashed line marks the width of the global $\pr(\kappax)$ for all
lines of sight.}
\label{fig:reconwidths}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[width=\columnwidth]{figs/Widthvshilberrt.eps}
\caption{Width of the \infered PDF $\pr(\kappax|\data)$ as a function of 
the true convergence, $\kappaxtrue$, for two different data qualities. 
Green assumes only photometric redshifts for all objects,
while red assumes a campaign of targeted spectroscopy.}
\label{fig:widthsvsH}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%

The best possible reconstructions of $\kappax$ will come from having a
spectroscopic redshift for every single object in each field.
\Fref{fig:reconwidths} shows the distribution of $\kappax$ posterior 
PDF widths for various data qualities: complete spectroscopic coverage results
in the smallest widths. In terms of telescope time, such a reconstruction
would likely be prohibitively expensive, but we investigate this scenario as
an ideal case.

\comment{PJM: The following does not belong here because \Fref{fig:widthsvsH}
does not show results for spectroscopic redshifts only!}

Reconstructing the lines of sight with perfect knowledge of the
redshift but an uncertain stellar mass, we find that the width of
$\pr(\kappax|\data)$ grows with the expectation of $\kappax$ and the ray-traced $\kappatrue$
(\Fref{fig:widthsvsH}). The low $\kappax$ lines of sight are relatively empty,
and so there are few opportunities for uncertainties in the halo masses to
\propogate into $\kappax$ uncertainties; low $\kappax$ lines of sight can be reconstructed
more precisely than high $\kappax$ lines of sight.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Reconstructing $\kappax$ from photometry alone}

Inferring the stellar mass of a galaxy from its magnitude and colours requires
an estimate of how far away the galaxy is; without a spectroscopic redshift
the \infered stellar mass is less precise. However, obtaining photometry has
much lower observational cost; all upcoming large area photometric surveys
will reach 24th magnitude, providing sufficient data to reconstruct lines of
sight  {\it without} additional observations. In principle the photometric
redshift is correlated with the \infered stellar mass, however we do not model
this effect since the convergence from the outskirts of an individual halo is
only weakly \dependant on redshift at fixed mass due to the breadth of the
lensing kernel: the redshift uncertainty has only a small effect on the
inferred $\pr(\kappax|\data)$ in comparison to that of  the uncertain stellar
masses.  With only photometric redshifts, the uncertainty on $\kappah$ is
larger than in the spectroscopic case, and this \propogates into a broader
$\pr(\kappax|\data)$, as can be seen in \Fref{fig:reconwidths}. However, the
photometric reconstruction still typically produces a 50\% improvement
compared to the precision of the global $\pr(\kappax)$.

\tom{With photometric data alone, $\kappah^{\rm med}$ can shift by $\sim$0.01
relative to $\kappah^{\rm med}$ given spectroscopic coverage. Since  the
$\kappax$ contribution of voids cannot change with data quality, the shift in
$\kappah^{\rm med}$ must be due to the asymmetric propagation of stellar mass
uncertainties into $\kappax$ uncertainties. As the stellar mass uncertainty
increases, $\kappah^{\rm med}$ is pushed higher.  To correctly calibrate 
$\pr(\kappax|\kappah^{\rm med},\data)$ one {\it must} include the data quality
used to generate $\kappah^{\rm med}$.}

We find that the width of $\pr(\kappax|\data)$ grows with the expectation
value of $\kappax$; this is shown in \Fref{fig:widthsvsH}. The low $\kappax$
lines of sight are relatively empty, and so there are few opportunities for
uncertainties in the halo masses to \propogate into $\kappax$ uncertainties.

%\flag{Is there a simple approximation for $\kappax$ and $\sigma_{\kappa}$ as a
%function of $\kappaxtrue$? This would be useful when making forecasts, writing
%proposals etc...}
%
%TEC: No. It's also not relevant to proposals since kappatrue isnt observable. One can ask is there
%a formula for sigmakappa given the expectation value of kappax. The answer to that is
%yes, but it depends on data quality. Concerning forecasting yes we would 
%want to use a simple approximation, but I don't think it has a place in this paper.
%When you/I/someone does the forecasting I'd like to be involved; I'll make the
%approximation then.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Reconstructing $\kappax$ with partial spectroscopic coverage}
\label{sec:obsMstar+z:targetedspec}

While a fully photometric reconstruction provides useful constraints on
$\pr(\kappax|\data)$, targeted spectroscopy can provide additional constraints
on the masses and redshifts of halos whose $\kappah$ contributions have the
largest absolute uncertainty. Obtaining spectra of bright ($i<22$) objects is
relatively fast with large telescopes: however, we find that if spectroscopic
redshifts were known for all $i<22$ galaxies in our fields the reconstruction
improves only slightly over the purely photometric reconstruction,  although
the improvement can depend strongly on the details of the particular line of
sight. Obtaining spectra for fainter objects would be correspondingly more
expensive, but if spectroscopic redshifts could be obtained for all $i<23$
galaxies (as part of a futuristic baryon acoustic oscillation survey, for
example) and all $i<24$ galaxies within 1 arcminute of the line of sight, then
we find that the $\pr(\kappax|\data)$ would be almost as precise as that from
having complete spectroscopic coverage (\Fref{fig:reconwidths}).
Consistent with our previous findings, the lowest $\kappax$ lines of
sight have the most constrained $\kappax$ PDFs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Systematic Errors due to Sample Selection}
\label{sec:biases}

\begin{figure*}
\includegraphics[width=\textwidth]{figs/biasplots.eps}
\caption{Accuracy in $\kappax$ from combining samples of lenses.  These plots
show the expectation value of $\prod_{i=1}^N \pr_i(\kappax-\kappaxtrue|\data)$ --
deviations from zero represent biases. The solid, dashed and dotted lines
correspond to combinations of  20, 5 and 2 sightlines respectively. Green
lines show the results \infered from photometry alone, whilst red lines show
the results  from the same targeted spectroscopy campaign described in 
\Sref{sec:obsMstar+z:targetedspec}. The three panels correspond to samples of
sightlines selected in different ways. Left: randomly selected lines of
sight.  Centre: sightlines randomly selected from the 33\% of lines whose
P($\kappax$) is most tightly constrained by our model.  Right: lines of sight
with external shear of 0.05 or greater.}
\label{fig:biasplots}
\end{figure*}

While it is good to reconstruct $\kappax$ precisely, it is more important that
$\kappax$ is reconstructed accurately. The reconstruction must not be biased. 
We quantify the accuracy of our reconstructed $\pr(\kappax|\data)$ in the
following way. By shifting the \infered $\pr(\kappax|\data)$ for each line of
sight by the true (ray-traced) convergence, we obtain a  PDF that should be
centred on zero; these offset PDFs can be multiplied together for multiple
lines of sight to emulate a joint likelihood analysis, and then test for
possible bias in it:
\be
\label{eq:bias}
\mathcal{P}_N = \prod_{i=1}^N \pr_i(\kappax-\kappax^{\rm true}|\data)
\ee
In this section, we quantify the bias as the size of the deviation in the
expectation value of $\mathcal{P}_N$ from zero. If the bias of $\mathcal{P}_N$
is smaller than the half-width of $\mathcal{P}_N$, our $\kappax$ inferences
can be considered accurate. Since the width of $\mathcal{P}_N$ decreases as
more sightlines are combined, a small bias will always 
eventually be found. In the context of measuring time-delay distance at the 
one percent level, the expectation value of $|\mathcal{P}_N|$ needs to be 
significantly less than 0.01. \tom{External convergence is not the only source
of uncertainty in measuring time-delay distance. Since the statistical
distance uncertainties due to lens modelling and time delay estimation are
likely to be at the 3-4\% level \citep{SuyuEtal2010}, convolving
$\mathcal{P}_N$ with a Gaussian of width $0.04/\sqrt{N}$ is a reasonable
approximation for the likely final uncertainty on time-delay distance.}

% \comment{From Sherry:
% Re: $P_N$:
% Not sure that this is the best way to quantify bias.  First of all, this
% doesn't include the modeling error of $\sim3-4\%$, so the number of lenses listed
% (for when <$P_N$> = half width of $P_N$) is underestimated.  Also, with this
% definition, a large number of lenses could be described as "biased" even if
% the resulting shift is say 0.0001, which won't really trouble us at all.  For
% example, what would the distributions look like in Fig.8 with 100 lenses? 
% Since the peaks of the red curves are within 0.01 of kappa_true, I would
% suspect that with a large number of lenses, the corresponding red curve would
% be within 0.01 of kappa_true.  The present day goal is to get H0 to 1\%, and
% it would be good to state somewhere that with targeted spectroscopy, one could
% recover kext at the 1\% level with a large sample of lenses.}

The source of systematic error that we investigate in this paper  is that of
sample selection: we define three different example selections of lines of
sight, and compute the bias  \phil{that would result (if left unaccounted
for)} in each one in turn, as a function of data quality. The results
described below are illustrated in \Fref{fig:biasplots}.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Selecting random lines of sight}
\label{sec:bias:random}

We first consider a purely random selection of lines of sight. With the purely
photometric reconstruction, there is no evidence of bias, even after combining
20 lines of sight. \tom{This result is a sanity check: if we had chosen lines
of sight with identical $\kappah^{\rm med}$ rather than $\kappah^{\rm
med}\pm0.003$ randomly selected lines of sight would have zero bias by
construction}. If lines of sight hosting time delay lenses are truely random, 
our method allows $\kappax$ to be reconstructed without inducing a bias (under
the basic assumptions that the universe is like the calibration simulation,
and the correct stellar-mass to halo-mass relation has been used).

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Selecting only the lines of sight with narrow $\pr(\kappax|\data)$}
\label{sec:bias:tightPDF}

Since follow-up campaigns (such as high resolution imaging, or time
variability monitoring) are expensive, it is likely that only a subset of
detected objects will be observed further. By pre-selecting objects that have 
the best-constrained convergence, the cosmological value per lens can be
increased. However, this selection \phil{(like any other)}
has the potential to induce a bias. It is
likely that only photometry will be available at the time of sub-sample
selection (although spectroscopic follow-up might be conducted at a later
date). We mimic such a selection by drawing lines of sight only from those in
the lowest third of $\sigma_{\kappa}$, given a purely photometric
reconstruction of their fields.

We find that selecting the most tightly constrained lines of sight in this way
does not introduce a significant bias:  our reconstruction is most accurate
for the lowest $\kappax$ lines of sight. Combining 20 such systems, the bias
is at the 0.001 level.

Photometry of the field seems to be sufficient to adequately select these
empty lines of sight.  However, 
since lenses are typically found in locally over-dense
environments, it is unlikely that  an almost empty line of sight would
actually  contain a lens; \phil{what we can say is that if the local
environment can be identified spectroscopically, and then the line
of sight mass distribution reconstructed using our method, selecting lenses
with low, well-constrained values of $\kappax$ will lead to an increase in
distance precision at no cost in bias.}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\subsection{Selecting only high shear lines of sight}
\label{sec:bias:tightPDF}

Time delay lenses are often selected for follow-up based on their images'
brightness (to make monitoring observations less expensive), their images'
separation (to decrease the covariance between the ground-based image
light-curves), or the fact that they are quadruply-imaging rather than
doubly-imaging (since this yields 3 time delays instead of 1, a higher
magnification and a more informative Einstein Ring system). All three of these
properties favour lenses with high convergence and shear along their lines of
sight. Focusing on the quad selection, we might expect many of these systems
to have significant external shear (since the cross-section for 4 image
production is so sensitive to ellipticity in the lens mass distribution (or
equivalently, external shear due to the lens' environment).

\phil{We can test the impact of such a shear selection by selecting only lines
of sight with an external shear above some threshold.  We define a somewhat
extreme selection  of $\gammax>0.05$, and find that in this sample, $\kappax$
would be systematically underestimated at the $\sim$0.008 level (corresponding
to a 0.8\% systematic error in distance).}  In the absence of other sources of
uncertainty, this systematic error would be significant in a sample of just 5
lenses; when including other time-delay distance uncertainties a 0.8\% bias would be
significant for $\sim$20 lenses. This systematic error is mitigated if only the lines of sight with both
$\gammax>0.05$ and well constrained $\pr(\kappax|\data)$ are used, but in this
case a bias at the $\sim$0.005 level remains.

While $\gammax>0.05$ is likely to be a much stronger selection than would
occur in reality, it is  nevertheless worth noting from this example that the
reconstruction \proceedure \phil{can be biased if lens selection functions are
extreme and unaccounted for.} Since our model does not include shear
constraints it is not surprising that a selection function based on shear can
induce a bias. A more sophisticated model that includes the shear recovered
from the lens modelling might be less susceptible to this bias.

The halo model can also be used to estimate the external shear along a line of
sight: shear is an observable that can be extracted from strong lens
modelling. However, there is a degeneracy between internal and external shear.
When the Einstein Ring imaging data are very good it is possible to
disentangle external and internal shear \citep[\eg][]{SuyuEtal2010}, but there
are still significant uncertainties. \citet{WongEtal2011} attempted to match
the shear from strong lens models with a reconstruction of the local lens
group environment, but found a tension between the strong lens model and the
reconstruction of the environment. Given the \citet{WongEtal2011} results, it
is unclear whether the external shear from lens models can be reconciled with
a line-of-sight reconstruction. Alternatively, it may be possible  to infer
external shear using weak lensing information from near the line of sight.  If
the true external shear can be measured, it provides an additional constraint
on which of the \MS lines of sight are similar to the reconstructed line of
sight. \citet{SuyuEtal2012} found that in the case of 
RXJ1131, combining shear constraints
with galaxy number count over-density gave a significantly different
$\pr(\kappax|\gamma,N_{45})$ compared to the PDF from number count
over-density alone, $\pr(\kappax|N_{45})$.

Extending our model to include shear we find that given
perfect knowledge of the halo mass and redshift the ray-traced external shear
$\gamma$ and the reconstructed external shear $\gamma_{\mathrm{h}}$  are
similar, with 68 percent of lines obeying
\be
\label{eq:shearineq}
|{\boldmath{\gammax}-\boldmath{\gamma_{\mathrm{h}}}}| < \mathrm{0.025}
\ee 
Future work should investigate whether $\gamma_{\mathrm{h}}$ can be used
to improve the accuracy and precision of $\kappax$ estimation, given a
reconstruction of the line of sight.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\section{Improving precision by including additional information}
%\notes{placeholder for shear}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}
\label{sec:discuss}

% \comments{
% We have shown that reconstructing the matter due to halos along any line of
% sight can improve the constraints on the external convergence along that line
% of sight, with, depending on sample selection,  the width of
% $\pr(\kappax|\data)$ being up to a factor of two smaller than the global
% $\pr(\kappax)$, and  we have quantified the residual bias in the $\kappax$
% estimates under our calibration assumptions. We now revisit these assumptions.
% }

The total convergence along a line of sight is strongly correlated with the
reconstructed $\kappah$. However, since our model ignores voids and assumes
all halos follow a spherical truncated-NFW profile our halo model does not
include all of the relevant physics; hence, the width of our resulting
$\pr(\kappax|\data)$ is still typically $\sim$0.01 for any given lightcone,
even assuming perfect knowledge of every halo's virial mass and redshift. To
make further progress a more advanced treatment of both voids and halos will
be needed. \citet{KarpenkaEtal2012} find that their supernova data prefer, in
the context of their simple halo model, a truncated isothermal profile for the
galaxy mass distributions. Galaxy-galaxy and cluster-galaxy  lensing studies
\citep[\eg][]{GavazziEtal2007,JohnstonEtal2007} also suggest an approximately
isothermal profile for the mass-shear correlation function; however, at large
radii these profiles are dominated by the effects of neighbouring halos, which
are already taken into account by our model.  Further iterations on the
analysis presented here should include the stellar mass distributions that
cause the isothermal density profiles in the centres of galaxies, although we
do not expect this to have a big impact on the reconstructions. Likewise, 
halo ellipticity may have some small effect on the predicted convergence. It
is possible that baryonic physics may alter    the radial profile of dark
matter halos, even on large scales. \citet{Semboloni+2012} found that baryonic
physics had a 30-40\% impact on three-point shear statistics, it is unclear
how baryons affect one-point convergence statistics compared to the pure dark
matter simulations used in this work.

Clusters, filaments and voids are more difficult to model, but simple forms
could in principle be derived from the structures found in the simulations,
and included in the halo model in some way.  Most importantly for the highest
$\kappax$ lines of sight will be a better group and cluster model:
\citet{MomchevaEtal2006} and  \citet{WongEtal2011} recommend using models
where dark mass is assigned to both galaxy-scale halos and halos for the
groups and clusters they occupy. To make this work, one could incorporate a
group and cluster finder, and then infer halo mass from optical richness
\citep[\eg][]{MaxBCG} in tandem with BCG stellar mass. \tom{Meanwhile,
improving the modelling of voids could be done using suggestions by
\citet{voids2} or \citet{voids1} combined with a void finder such as in
\citet{voids3}. Combining an advanced halo model with a void model will allow
a direct test of the halo approximation: using real photometric data one could
reconstruct lines of sight and compare against measured weak lensing shears.
This would ensure that the reconstruction model reproduces the contents of the
real universe, rather than a particular simulation.}

% \comments{Interestingly, we have found that it is the most empty lines of sight that 
% can be reconstructed with the highest precision. $\kappax$ estimates for empty
% lines of sight receive few contributions from halos, and a large contribution
% from voids; since they have the tightest PDFs after the reconstruction it
% seems that our model's biggest uncertainties are driven by naively
% reconstructing halos rather than neglecting voids. With a photometric
% reconstruction of the field there is a significant broadening, but the
% resulting precision is still typically 50\%  higher than using the global
% $\kappax$ distribution.
% 
% Since the uncertainty on $\pr(\kappax|\data)$ is $\sim$0.01 even
% with perfect knowledge of halo mass and redshift, it seems that modelling uncertainties
% from voids, deviations from sphericity and dark matter clumping within the
% main halo dominate the total $\kappax$ uncertainty even with very
% uncertain stellar masses. Inferring halo ellipticity and dark matter
% clumping will likely always remain a difficulty for line of sight
% reconstruction; but as \Fref{fig:magcut} shows, observing deeper
% than 24th magnitude does not significantly help the reconstruction.
% }

% \comments{For a small fraction of lines of sight, $\pr(\kappax|\data)$ remains very
% broad even after applying our reconstruction; these lines of sight are
% typically the most over-dense lines of sight in the universe. For
% time-delay cosmography the most observationally expensive tasks are the
% light-curve monitoring and high resolution follow-up imaging and spectroscopy,
% while making photometric observations of a
% 4$\times$4 arcminute patch of sky survey down to 24th magnitude is
% relatively cheap. We have shown that with a single epoch observation of
% the region around a strong lens it is possible to infer $\pr(\kappax|\data)$.
% If a line of sight has a broad $\pr(\kappax|\data)$ it can be rejected {\it
% before} the investment of follow-up. We find that
% selecting only the lines of sight with the most tightly constrained
% $\pr(\kappax)$ does not induce a bias in $\kappax$, and hence distance.
% For the most over-dense lines of sight,
% the precision of $\pr(\kappax|\data)$ is improved noticeably by the addition
% of spectroscopic information; currently insufficient time-delay systems are
% know to choose a subset based on their reconstructed $\pr(\kappax|\data)$.
% Spectroscopic coverage will definitely be useful for extracting cosmology from
% over-dense lines of sight.
% }
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \comments{%
% \begin{figure}
% \includegraphics[width=\columnwidth]{figs/biasplots3.eps}
% \caption{Convergence estimation accuracy using the full  $\pr(\kappax|\kappah)$
% distribution, rather than its median.  As in \Fref{fig:biasplots}, solid,
% dashed and dotted lines show combinations of 20, 5 and 2 lines of sight
% respectively, while green lines show the inferences 
% from photometry alone, red lines from photometry and targeted spectroscopy 
% (assuming a spectroscopic redshift for all
% halos with $i<23$ and all halos within 1 arcminute and $i<24$), and blue lines
% from the unrealistic assumption of a spectroscopic redshift for all galaxies
% in the field. }
% \label{fig:no-median-tricks}
% \end{figure}
% }%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The reduction of $\pr(\kappah|\data)$ to just its median value, before using
it to look up the appropriate $\pr(\kappax)$ from the simulated sightline
ensemble, is reminiscent of the procedure followed by \citet{SuyuEtal2010},
and Greene et al (submitted). The reconstruction method presented here can be
seen as the limiting case of the weightings explored by Greene et al: rather
than using weighted number counts in an aperture, we treat each galaxy
individually, and predict $\kappa_{\rm h,obs}^{\rm med}$ for use in their
place. Given photometry alone, our reconstruction represents a $\sim$50\% improvement of
the precision compared the number counts used in \citet{SuyuEtal2010}. Greene et al.
made important progress for the most over-dense lines of sight; our
reconstructions are $\sim$30\% more precise than those of Greene et al.

% \comments{ One
% significant disadvantage of these types of method is that they impose a
% bottleneck in the flow of information. Once you have compressed to one number,
% it is difficult to incorporate more information later; you have to recompute
% all the calibrations. Using the full PDF for $\kappah$ is the more correct
% thing to do: at present, however, it suffers from significant bias in the
% photometric data case. \Fref{fig:no-median-tricks} shows how $\kappax$ is
% overestimated, as a result of the high scatter in halo mass for objects with 
% high stellar mass. As has already been noted, 
% future work should focus on increasing the precision of the halo
% mass estimates. 
% }

In this work, we have used  the \MS as a calibration tool. If the real
Universe is not like the \MS, then this calibration will be incorrect.
Repeating this  study using a different simulation to make the test catalogues
would allow us to quantify the sensitivity of our results to the simulation
used. Similarly the use of a different semi-analytic model to paint galaxies
onto halos, and an alternative stellar mass--halo mass relation would both
enable similar exploration of these systematic errors. \citet{SuyuEtal2010}
used the relative over-density to mitigate this effect, but it does not
completely remove the dependence on simulation. Future work should include
investigating the robustness of the reconstruction to these systematic errors.

% \comments{In particular, the method presented here is tied to the simulations at two
% points: the prior PDF for $\kappah$, and the conditional distribution
% $\pr(\kappax|\kappah)$ used to convert halo model convergence to total
% convergence. The first tie could potentially be avoided if the halo mass
% estimation gave rise to a reasonable $\pr(\kappah|\data)$ on its own.
% Possibilities for improving the precision of the halo mass estimation are
% discussed above. The second tie is more difficult, since breaking it means
% modelling the unseen mass structures and voids. It
% should be possible to infer hierarchically  the parameters of a flexible model
% for these things from the data itself, as demonstrated by
% \citet{KarpenkaEtal2012} (albeit under the assumption of scatter-free scaling
% relations) -- how this affects the cosmographic precision is to be
% investigated. When working with wide field imaging surveys,  weak lensing shears will be
% available around every line of sight:  fitting these with the halo models
% described here in order to constrain the hyper-parameters of the model presents a 
% possible test of our halo model prescription against real observations.
% }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:conclude}

In this work we have investigated a simple halo model prescription for
reconstructing all the mass along a line of sight to an intermediate redshift
source. We have used the ray-traced lensing convergence along lines of sight
through the Millennium Simulation to calibrate estimates of the total
convergence made by summing the convergences due to each object in a
photometric catalogue. Having found that the reconstruction process is
accurate given this calibration and perfect knowledge of halo mass and
redshift, we investigated the effects of reasonable uncertainties in the
stellar mass and redshift of each halo, and propagated these uncertainties
into a $\pr(\kappax|\data)$ for each line of sight. We also defined three
different possible line of sight selections, and investigated the possible 
bias induced by these selections. We draw the following conclusions:

\begin{itemize}

\item Despite our model's simplicity, its reconstructed $\kappah$ values are
good tracers of $\kappax$. $\kappah$ is biased due to our ignorance of voids,
groups and clusters, and the unseen mass in small halos and filaments, but we
find that this can be calibrated out using the Millenium Simulation
catalogues. We found that with perfect knowledge of every halo's mass and
redshift,  calibrated reconstruction of a typical line of sight  gives an
unbiased estimate of $\kappax$ that is not perfect, but is $1.8$ 
times more precise than the global
$\pr(\kappax)$. This factor quantifies the limit to which the halo model can
represent the external convergence due to line of sight mass structure.

\item With uncertain halo masses and redshifts, we find that
$\pr(\kappah|\data)$ can still be calibrated in order to infer
$\pr(\kappax|\data)$ from the ensemble of simulated lines of sight, but the
resulting PDFs tend to be much broader than for perfect halo mass and redshift
reconstructions.

\item It is very rare for halos further than 2 arcminutes from the line of
sight to make a significant contribution to $\pr(\kappax|\data)$; likewise, 
including
halos whose host galaxy is less luminous than $i=24$ does not significantly
improve our reconstructions.  A photometric survey to this depth of a
4$\times$4 arcminute patch around the target would approach the limiting
uncertainties of our simple reconstruction recipe, and yield a 
$\pr(\kappax|\data)$ that has, on average, a width of 0.016 (inducing a $\sim$1.6\%
uncertainty on time-delay distances). This is 1.5 times
less broad than the global $\pr(\kappax)$.

\item  For the most over-dense lines of sight, the reconstruction produces a
broad PDF; conversely, we find that the lines of sight with the sharpest
inferred $\pr(\kappax|\data)$ are typically under-dense. Since the
reconstructions described here can be performed before follow-up time is
invested, it will be possible to select targets by the precision of their
$\kappax$ estimates. This will be useful in an era when there are more known
lenses than can be followed-up.

\item  Photometric data is sufficient to select the best lines of sight for
follow-up. Spectroscopic coverage helps to tighten  $\pr(\kappax|\data)$ for
high $\kappax$ lines of sight.
% , but is of less use for low $\kappax$ lines of
% sight.

\item Selecting from the highest precision 33\% of the reconstructed lines of
sight results in a sample with low bias, $\sim$0.1\% in terms of the
time-delay distance. \tom{For these lines of sight the reconstructed
$\pr(\kappax|\data)$ induces a statistical uncertainty of $\sim$1\% on
individual time-delay distances.}

\item Conversely, selecting lines of sight with high shear ($|\gamma| > 0.05$
(a somewhat extreme selection, but one that  could potentially arise from 
focusing on four-image or wide separation lenses) results in a sample with
bias of 0.8\% in time delay distance. Including realistic estimates of other
sources of uncertainty a systematic error of 0.8\% would become significant in
a sample of $\sim$20 lenses.

% \item With only a photometric reconstruction of all time delay lens fields,
% and following up only the systems with the most precise $\kappax$ estimates,
% the width of the dominant statistical uncertainty in time-delay cosmography
% can be halved, without inducing biases at the one percent level.

\end{itemize}

% PJM: Im not sure this is the right note to end on.

% TEC: I liked it! but it's open to discussion.

% External convergence is not the only uncertainty in time delay cosmology, but
% for systems like B1608$+$656 it is the dominant one. For most lines of sight
% lens modelling induces uncertainties at the $\sim$4 percent
% level, which far exceeds the typical $\kappax$ uncertainties after applying our
% reconstruction \proceedure. By using reconstruction methods such as that
% explored here to select the most appropriate 
% lenses for follow-up, the uncertainties due to line of sight
% convergence will become a sub-dominant source of error.

While we have tested our reconstruction method against realistic simulations
of line of sight mass distributions in the Universe, our method is also 
calibrated to these simulations. Testing this assumption in the short term,
and replacing it with an empirical calibration (or hierarchical inference) in
the longer term, are worthy goals for future work. The halo model framework is
flexible enough to enable many such improvements, including, quite naturally,
the incorporation of more information from observations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  ACKNOWLEDGMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Acknowledgements}

TEC thanks Vasily Belokurov for supervision, guidance and suggestions.
We thank Risa Wechsler and Peter Behroozi 
for useful discussions and suggestions.
\input{acknowledgments.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  APPENDICES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Inferring $\Mhalo$ given a noisy measurement of $\Mstarobs$}
\label{appendix:MSMH}

To estimate the convergence caused by a halo we need to know its mass; how can 
halo mass be \infered from a noisy estimate of the stellar mass $\Mstarobs$
of a galaxy at redshift $z$. We seek the posterior
PDF $\Pr(\Mhalo|\Mstarobs,z)$, which can be expanded as follows:

\begin{eqnarray}
&& \Pr(\Mhalo|\Mstarobs,z) = \notag\\
&& \int d\Mstar \Pr(\Mhalo|\Mstar,z) \Pr(\Mstar|\Mstarobs,z), \notag\\
&\propto& \int d\Mstar \Pr(\Mstarobs|\Mstar) \Pr(\Mstar|\Mhalo,z) \Pr(\Mhalo|z),
\label{eq:mhalo-mstar}
\end{eqnarray}
where we have used Bayes' Theorem twice to replace
$\Pr(\Mhalo|\Mstar,z) \Pr(\Mstar|z)$ with 
$\Pr(\Mstar|\Mhalo,z) \Pr(\Mhalo|z)$, and 
to invert $\Pr(\Mstar|\Mstarobs)$ into the sampling
distribution $\Pr(\Mstar|\Mstarobs)$, which we recognise as the likelihood
function for the observed stellar mass. Note that the ``true'' $\Mstar$ of the
galaxy is marginalised out: we are only interested in inferring the halo
mass. The last two terms in
\eqref{eq:mhalo-mstar} are the $\Mstar-\Mhalo$ relation from
\citet{BehrooziEtal2010}, and the halo mass function $\Pr(\Mhalo|z)$, at the
given redshift. We can
tabulate the product of these two from our Millennium Simulation catalogue,
constructing a two-dimensional histogram of halo masses and their associated
true stellar masses (drawn from the Behroozi relation).

For each galaxy, we compute the likelihood function for its $\Mstarobs$ as a
function of the unknown $\Mstar$, and multiply it by our tabulated joint PDF.
This heavily downweights halos with $\Mstar$ values outside the observed
range. We then do the marginalisation integral by Monte Carlo, drawing
(two-dimensional) sample parameter vectors
from the down weighted histogram, discarding the $\Mstar$ values, and
constructing a one-dimensional histogram that is an estimate of
$\Pr(\Mhalo|\Mstarobs)$.

If the redshift of the galaxy is uncertain, we need to take this
uncertainty into account; for example, for each sample drawn from the
photometric redshift posterior PDF $\Pr(z|{\rm colors})$, we can draw a
sample $\Mhalo$ using the above procedure.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  REFERENCES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% MNRAS does not use bibtex, input .bbl file instead. 
% Generate this in the makefile using bubble script in scriptutils:

% bubble -f paper-lcr.tex references.bib 
% \input{paper-lcr.bbl}

% \bibliographystyle{apj}
% \bibliography{references}

\input{references.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\label{lastpage}
\bsp

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%